---
title: "bothClust simulation results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads and plots results from the standalone "bothClust" simulation scripts
(`source/simulations/k_bothClust_expectation.R` and `source/simulations/k_bothClust_variance.R`),
addressing the reviewer request for a scenario with true clustering in both immune and
background cells. Two variants are compared:

- `bothClust`: immune and background clustered via **independent** kernels (separate
  cluster locations).
- `bothClustDep`: immune kernel **dependent** on (correlated with) the background
  kernel (overlapping/coincident clusters) - a harder scenario where the immune-specific
  signal is confounded with local background density.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(spatstat.random)
library(patchwork)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

my_colors <- c(
  "K"         = "#6A3D9A",  # purple
  "Kinhom"    = "#1F78B4",  # blue
  "perm"      = "#009E73",  # bluish green
  "KAMP"      = "#D55E00",  # vermillion
  "KAMP lite" = "#E69F00"   # orange
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```


# Expectation: degree of clustering

Loads results from `output/bothClust_expectation/`, produced by
`k_bothClust_expectation.R`. Same "degree of clustering" statistic
(`trans - csr`) used in `k_expectation.Rmd`'s `pBothClust` figure, computed
here from the standalone bothClust-only run.

```{r}
files_exp = list.files(here::here("output", "bothClust_expectation"),
                       full.names = TRUE, recursive = TRUE, pattern = ".RDA")

df_exp = map_dfr(files_exp, merge_files)
```

```{r}
df_exp = df_exp %>%
  filter(r == 1.25,
         abundance != 0.01) %>%
  mutate(abundance = factor(abundance),
         value = trans - csr,
         method = factor(method,
                         levels = c("k", "kinhom", "perm",
                                    "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP",
                                    "KAMP lite")),
         type = factor(type, levels = c("bothClust", "bothClustDep"),
                      labels = c("independent clusters", "dependent clusters")),
         lambda_n = str_c("lambda[n]: ", lambda_n),
         lambda_n = factor(lambda_n, levels = c("lambda[n]: 1000", "lambda[n]: 2000",
                                                "lambda[n]: 5000", "lambda[n]: 10000")))

# trim extreme outliers dynamically (rather than a hardcoded cutoff) since
# combinatorics in get_permutation_distribution() can blow up at very low
# immune cell counts
bounds_exp = quantile(df_exp$value, c(0.01, 0.99), na.rm = TRUE)

p_bothClust_expectation = df_exp %>%
  filter(value >= bounds_exp[1], value <= bounds_exp[2]) %>%
  ggplot(aes(abundance, value, group = interaction(abundance, method), fill = method)) +
  geom_boxplot() +
  labs(x = "Proportion of immune cells",
       y = "Degree of clustering",
       title = "Alternative: background and immune both clustered") +
  theme_minimal() +
  scale_fill_manual(values = my_colors) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray") +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  facet_grid(type ~ lambda_n, labeller = labeller(lambda_n = label_parsed, type = label_value))

p_bothClust_expectation
```


# Variance: power

Loads results from `output/bothClust_variance/`, produced by
`k_bothClust_variance.R`. Because immune cells are genuinely clustered
(distinct from, in the independent case, or confounded with, in the
dependent case, background clustering), this is a true alternative-hypothesis
scenario: Power = P(reject H0 | H1 true).

```{r}
files_var = list.files(here::here("output", "bothClust_variance"),
                       full.names = TRUE, recursive = TRUE, pattern = ".RDA")

df_var = map_dfr(files_var, merge_files, variance = TRUE) %>%
  filter(method != "kperm approx") %>%
  mutate(method = factor(method, levels = c("kperm", "kamp", "kamplite"),
                         labels = c("perm", "KAMP", "KAMP lite")),
         type = factor(type, levels = c("bothClust", "bothClustDep"),
                      labels = c("independent clusters", "dependent clusters")))
```

```{r}
power_df = df_var %>%
  filter(r == 0.25) %>%
  group_by(type, method, lambda_n, abundance) %>%
  summarize(power = sum(pvalue <= 0.05, na.rm = TRUE) / sum(!is.na(pvalue)),
            n_sim = n(),
            .groups = "drop") %>%
  mutate(abundance = factor(abundance),
         lambda_n = factor(lambda_n))

power_df %>%
  arrange(n_sim)

p_bothClust_power = power_df %>%
  ggplot(aes(lambda_n, power, group = method,
             color = method, linetype = method)) +
  geom_point(size = 3) +
  scale_color_manual(values = my_colors) +
  geom_line(linewidth = 1.25) +
  labs(y = "Power", x = expression(lambda[n]),
       title = "Background and immune both clustered") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.9, linetype = 4) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  facet_grid(type ~ abundance, labeller = labeller(abundance = label_both, type = label_value))

p_bothClust_power
```


```{r fig_sim_bothClust_combined, fig.width=10, fig.height=8}
p_bothClust_expectation / p_bothClust_power
```
