---
title: "simulation results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the paper. Libraries and code are sourced below. 

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(spatstat.random)
library(patchwork)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)


my_colors <- c(
  "K" = "#440154",
  "Kinhom" = "#414487",
  "perm" = "#2A788E",
  "KAMP" = "#7AD151",
  "KAMP lite" = "#FDE725"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```


# Expectation

## Clean and process simulation data

Here we load and process simulation results. Full simulations involve 36 scenarios with 50 simulated datasets each.


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "univariate_survival", "20241118"), 
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)

```




```{r}

df = df %>%
    mutate(method = factor(method,
                         levels = c("k", "kinhom", "perm", 
                                    "kamp",  
                                    "kamp_lite"),
                         labels = c("K", "Kinhom","perm", "KAMP",
                                    "KAMP lite")),
           n_subj = factor(n_subj),
           term = ifelse(term %in% c("k", "kamp"), "doc", term)) 
```



To calculate

- mean squared error across true beta values
- power and type 1 error 
- might want to think about mimicking what the distribution of the measurement error in the real data looks like
- also want to look at bias in X2 and make x2 have a similar magnitude, direction, and correlation as in the real data

- Save n_subj, save values of other betas, add in a beta that is not correlated but associated
_ Check if mean of K is in the right direction
- Kamp seems to always be zero? Maybe my code is set up weird.  Add Normal distributed errors to KAMP directly 

```{r}

df %>%
  group_by(method) %>%
  mutate(bias = case_when(term == "doc" ~ estimate - beta_val,
                          term == "X2" ~ estimate + 1,
                          TRUE ~ estimate - 1)) %>%
  ungroup() %>%
  filter(#rho == -0.5, 
         n_subj == 500) %>%
  mutate(beta = factor(beta_val)) %>%
  ggplot(aes(beta, bias, group = interaction(beta, method), fill = method)) +
  geom_boxplot() +
  labs(x = "beta true",
       y = "bias") +
  theme_minimal() +
  scale_fill_manual(values = my_colors) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray") +
  theme(legend.position = c(0.6, 0.2),
        legend.direction = "horizontal",
        legend.title = element_blank()) +
  #ylim(-3,3) +
  facet_grid(rho~term, labeller = label_both) 

```


Power


```{r}
df %>%
  filter(!is.na(p.value),
          beta_val != 0, term == "doc") %>%
  group_by(term, method, n_subj, rho, beta_val) %>%
  summarize(power = sum(p.value<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
  ggplot(aes(n_subj, power, group = method,
             color = method, linetype = method)) +
  geom_point(size = 2) +
  scale_color_manual(values = my_colors) +
  geom_line(linewidth = 1.25) +
  #labs(y = "Type 1 error") +
  geom_hline(yintercept = 0.05, linetype = 4) +
  #theme(legend.position = "none") +
  facet_grid(beta_val~rho,  labeller = label_both)
```


Type 1 error

```{r}
df %>%
  filter(!is.na(p.value),
         beta_val == 0, term == "doc") %>%
  group_by(term, method, n_subj, rho, beta_val) %>%
  summarize(type1 = sum(p.value<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
  ggplot(aes(n_subj, type1, group = method,
             color = method, linetype = method)) +
  geom_point(size = 2) +
  scale_color_manual(values = my_colors) +
  geom_line(linewidth = 1.25) +
  labs(y = "Type 1 error") +
  geom_hline(yintercept = 0.05, linetype = 4) +
  #theme(legend.position = "none") +
  facet_wrap(~rho, ncol = 3, labeller = label_both)

```



## Computation time


```{r}
p_times = times_df %>%
  filter(type == "hom") %>%
  mutate(abundance = factor(abundance),
         lambda_n = factor(lambda_n,
                           labels = c("lambda[n]: 1000",
                                      "lambda[n]: 2000",
                                      "lambda[n]: 5000",
                                      "lambda[n]: 10000"))) %>%
  ggplot(aes(abundance, log(time), color = method, 
             group = interaction(abundance, method), linetype = method)) +
  geom_point() +
  theme_minimal() + 
  theme(legend.position = "top",
        legend.title = element_blank()) +
  geom_line(aes(group = method), linewidth = 1.25) +
  geom_hline(yintercept = log(1), linetype = 4, color = "gray") +
  scale_color_manual(values = my_colors) +
  labs(x = "abundance",
       y = "median log(seconds)") +
  facet_wrap(~lambda_n, labeller = label_parsed, nrow = 1)
  
```





```{r fig_sim_expectation2, fig.width=10, fig.height=12}

(p_hom+p_inhom)/(phomClust+pinhomClust)/p_times

```


```{r fig_sim_expectation, fig.width=10, fig.height=8}

(p_hom+p_inhom)/(phomClust+pinhomClust)

```

```{r fig_sim_comp, fig.height=3}

p_times

```



