---
title: "simulation results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the paper. Libraries and code are sourced below. 

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(spatstat.random)
library(patchwork)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)


my_colors <- c(
  "K" = "#440154",
  "Kinhom" = "#414487",
  "perm" = "#2A788E",
  "KAMP" = "#7AD151",
  "KAMP lite" = "#FDE725"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```


# Expectation

## Clean and process simulation data

Here we load and process simulation results. Full simulations involve 36 scenarios with 50 simulated datasets each.


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "univariate_survival", "20241120"), 
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)

```




```{r}

df = df %>%
    mutate(method = factor(method,
                         levels = c("k", "kinhom", "perm", 
                                    "kamp",  
                                    "kamp_lite"),
                         labels = c("K", "Kinhom","perm", "KAMP",
                                    "KAMP lite")),
           n_subj = factor(n_subj),
           term = ifelse(term %in% c("k", "kamp"), "DoC", term),
           beta_term = str_c("beta: ", term)) %>%
    group_by(method) %>%
    mutate(bias = case_when(term == "DoC" ~ estimate - beta_val,
                            term == "X2" ~ estimate + 1,
                          TRUE ~ estimate - 1)) %>%
    ungroup() 
```



# Plots 

Below shows the bias as a boxplot Figure

```{r }
bias = df %>%
  filter(n_subj == 500) %>%
  filter(term == "DoC") %>%
  mutate(beta = factor(beta_val)) %>%
  ggplot(aes(beta, bias, group = interaction(beta, method), fill = method)) +
  geom_boxplot() +
  labs(x = TeX("True $\\beta$ for degree of clustering"),
       y = TeX("$\\hat{\\beta}-\\beta$"),
       title = TeX("Bias")) +
  theme_minimal() +
  ylim(-3,3) +
  scale_fill_manual(values = my_colors) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray") +
  theme(legend.position = c(0.6, 0.8),
        legend.direction = "horizontal",
        legend.title = element_blank(),
         legend.text = element_text(size = 12)) 


bias
```


Coverage as a Figure


```{r}
coverage_df = df %>%
  filter(!is.na(conf.low)) %>%
  dplyr::select(term, rho, method, beta_val, n_subj, conf.high, conf.low) %>%
  filter(term == "DoC",
         abs(beta_val) < 2) %>%
  group_by(method, beta_val, n_subj) %>%
  mutate(coverage = ifelse(beta_val < conf.high & beta_val > conf.low, 1, 0)) %>%
  summarize(coverage = sum(coverage),
            n_sim = n()) %>%
  ungroup() %>%
  mutate(lower = map2_dbl(coverage, n_sim, ~ {binom.test(.x, .y)$conf.int[1]}),
         upper = map2_dbl(coverage, n_sim, ~ {binom.test(.x, .y)$conf.int[2]}),
         coverage = coverage / n_sim) %>%
  mutate(beta_val = str_c("beta: ", beta_val),
         beta_val = factor(beta_val, levels = c("beta: -2", "beta: -0.5", "beta: 0", "beta: 0.5", "beta: 2")))


cov = coverage_df %>%
  ggplot(aes(n_subj, coverage, group = method, color = method)) +
  geom_point(position = position_dodge(width = 0.5), size = 2.5) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.3,
                position = position_dodge(width = 0.5)) +
  labs(x = "number of subjects",
       title = "Coverage") +
  theme(legend.position = c(0.5, 0.4),
         legend.title = element_blank(),
         legend.text = element_text(size = 12)) + 
  scale_color_manual(values = my_colors) +
  geom_hline(yintercept = 0.95, linetype = 2, color = "gray") +
  facet_wrap(~beta_val, labeller = label_parsed, nrow = 1)

cov
```






```{r fig_sim_bias, fig.height = 5}
bias + cov + plot_layout(widths = c(1,1.5))
```







Power


```{r, eval = FALSE}
df %>%
  filter(!is.na(p.value),
          beta_val != 0, term == "doc") %>%
  group_by(term, method, n_subj, rho, beta_val) %>%
  summarize(power = sum(p.value<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
  ggplot(aes(n_subj, power, group = method,
             color = method, linetype = method)) +
  geom_point(size = 2) +
  scale_color_manual(values = my_colors) +
  geom_line(linewidth = 1.25) +
  #labs(y = "Type 1 error") +
  geom_hline(yintercept = 0.9, linetype = 4) +
  #theme(legend.position = "none") +
  facet_grid(beta_val~rho,  labeller = label_both)
```


Type 1 error

```{r, eval = FALSE}
df %>%
  filter(!is.na(p.value),
         beta_val == 0, term == "doc") %>%
  group_by(term, method, n_subj, rho, beta_val) %>%
  summarize(type1 = sum(p.value<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
  ggplot(aes(n_subj, type1, group = method,
             color = method, linetype = method)) +
  geom_point(size = 2) +
  scale_color_manual(values = my_colors) +
  geom_line(linewidth = 1.25) +
  labs(y = "Type 1 error") +
  geom_hline(yintercept = 0.05, linetype = 4) +
  #theme(legend.position = "none") +
  facet_wrap(~rho, ncol = 3, labeller = label_both)

```


