---
title: "simulation results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the paper. Libraries and code are sourced below. 

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(spatstat.random)
library(patchwork)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```



# Process simulation data

Here we load and process simulation results. Full simulations involve 18 scenarios with 100 simulated datasets each.


```{r}

# load only results for VB and TS
files = list.files(here::here("output", "simulation_results"), 
									 full.names =  TRUE)

# load all results
results = c()
#results_mse = vector("list", length = length(files))
for(f in 1:length(files)){
	load(files[f])
	results = rbind(results, k_univariate)
}

df = as_tibble(results) 

rm(f, files, k_univariate, results)
```


Do some data cleaning

```{r}

df = df %>%
  mutate(type_char = case_when(
    type == 1 ~ "hom",
    type == 2 ~ "inhom",
    type == 3 ~ "homClust",
    TRUE ~ "inhomClust"
  )) %>%
    mutate(k_theo = khat - ktheo,
         k_inhom = kinhomhat - kinhomtheo,
         k_perm = khat - kperm,
         k_fperm = khat - kfperm) %>%
  select(r, scenario, type_char, n, nm, lambda_n, lambda_nm, 
         contains("k_"), contains("time")) 

```


# Plots


* mise_bord gives a sense of what amount of clustering is actually happening (difference between CSR and observed K)
* Kinhom - is there a CSR K inhom value?  Yes- same us for homo. Compared Kinhom to homo. This should only work well in scenarios when there is not clustering.



Maybe the "new statistic" should be difference between observed and permuted, and should be compared with Kinhom? I'm guessing it will work better than Kinhom when data is clustered.   


## All


```{r 2024_simClust}
df %>%
  filter(lambda_nm == 50, r == 0.1) %>%
  pivot_longer(starts_with("k_"), names_to = "method", values_to = "value") %>%
  mutate(method = str_remove(method, "k_"),
         lambda_n = factor(lambda_n),
         lambda_nm = factor(lambda_nm),
         type_char = factor(type_char, levels = c("hom", "inhom", "homClust", "inhomClust"),
                            labels = c("Homogeneous, no clustering", "Inhomogeneous, no clustering", 
                                       "Homogeneous, clustering", "Inhomogeneous, clustering")),
         method = factor(method, levels = c("theo", "perm", "fperm", "fpermThin",
                                            "inhom"),
                         labels = c("theo", "perm", "fperm", "thin",
                                    "inhom"))) %>%
  #filter(type_char != "inhomClust") %>%
  ggplot(aes(method, value, group = method, fill = method)) +
  geom_boxplot()  +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "", y = "Degree of clustering") +
  ylim(-.05, .4) +
  theme(legend.position = "none") +
  facet_wrap(~type_char, ncol = 4)
```


```{r 2024_simTime}
time_df = df %>%
  filter(r == 0.1) %>%
  pivot_longer(contains("time_"), names_to = "method", values_to = "value") %>%
  mutate(method = str_remove(method, "time_"),
         lambda_n = factor(lambda_n),
         lambda_nm = factor(lambda_nm),
         type_char = factor(type_char, levels = c("hom", "inhom", "homClust", "inhomClust"),
                            labels = c("Homogeneous, no clustering", "Inhomogeneous, no clustering", 
                                       "Homogeneous, clustering", "Inhomogeneous, clustering")),
         method = factor(method, levels = c("k", "kperm", "fperm", "fperm_thin", "kinhom"),
                         labels = c("theo", "perm", "fperm", "thin", "inhom"))
         ) %>%
  filter(!is.na(method))


time_df %>%
  filter(type_char == "Inhomogeneous, clustering") %>%
  ggplot(aes(method, log(value), group = method, fill = method)) +
  geom_boxplot()  +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "", y = "time (seconds)") +
  theme(legend.position = "none") +
  facet_wrap(lambda_nm~lambda_n, scales = "free")
```

