---
title: "simulation results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the paper. Libraries and code are sourced below. 

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(spatstat.random)
library(patchwork)
library(ggbreak)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```


# Expectation

## Clean and process simulation data

Here we load and process simulation results. Full simulations involve 36 scenarios with 50 simulated datasets each.


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "univariate_expectation", 
                              "varyAbundance", "20240929"), 
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)


files = list.files(here::here("output", "univariate_expectation", 
                              "varyAbundance_nperm100", "20240929"), 
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df_perm100 = map_dfr(files, merge_files)
```




```{r}
# make comp times dataset
# ignoring holes datasets since we don't need all those scenarios
times_df = df %>%
  select(r, iter, type = type_char, lambda_n,
         abundance,
         contains("time_")) %>%
  filter(r == .15) %>%
  pivot_longer(time_kinhom:time_kperm, 
               names_to = "method", values_to = "time",
               names_prefix = "time_") %>%
  bind_rows(., mutate(df_perm100, method = "perm100") %>%
              select(r, iter, type = type_char,
                      lambda_n, abundance, method, 
                     time = time_kperm)) %>%
  mutate(method = factor(method,
                         levels = c("k", "kinhom", "perm100", 
                                    "kperm",  
                                    "kepd", "kepdThin"),
                         labels = c("K", "Kinhom","perm100", "perm1000", "KAMP",
                                    "KAMP lite"))) %>%
  group_by(method, type, lambda_n, abundance) %>%
  summarize(time = median(time)) %>%
  ungroup() 

# degree of clustering
doc_df = df %>%
  select(r, iter, type = type_char, lambda_n,
         abundance,
         starts_with("k")) %>%
  filter(r == .15) %>%
  pivot_longer(ktheo:kepdThin, 
               names_to = "method", values_to = "value") %>%  
  bind_rows(., mutate(df_perm100, method = "perm100") %>%
              select(r, iter, type = type_char,
                      lambda_n, abundance, method, 
                     value = kperm)) %>%
  pivot_wider(names_from = method, values_from = value) %>%
  mutate(k = khat - ktheo,
         inhom = kinhomhat - kinhomtheo,
         perm100 = khat - perm100,
         perm1000 = khat - kperm,
         epd = khat - kepd,
         epdThin = khat - kepdThin) %>%
  select(r, iter, type, lambda_n, abundance, k, inhom, perm100,
         perm1000, epd, epdThin) %>%
  pivot_longer(k:epdThin, names_to = "method", values_to = "value") %>%
  mutate(method = factor(method, levels = c("k", "inhom", "perm100",
                                            "perm1000", 
                                            "epd", "epdThin"),
                         labels = c("K", "inhom", "perm100", "perm1000", 
                                    "KAMP", "KAMP lite")),
         type_char = factor(type, levels = c("hom", "inhom", "homClust", "inhomClust"),
                            labels = c("homogeneous",
                                       "inhomogeneous",
                                       "clustered",
                                       "inhomogeneous + clustered")))
```





## Degree of clustering


```{r}
doc_df %>%
  filter(lambda_n == 5000) %>%
  mutate(abundance = factor(abundance))  %>%
  ggplot(aes(abundance, value, group = interaction(abundance, method), fill = method)) +
  geom_boxplot() +
  labs(x = "abundance",
       y = "degree of clustering") +
  theme_minimal() +
  theme(legend.position = "none",
        legend.title = element_blank()) +
  facet_wrap(~type, scales = "free", nrow = 1)
```






## Computation time


```{r}
p_times = times_df %>%
  filter(type == "inhom", lambda_n != 100) %>%
  mutate(abundance = factor(abundance),
         lambda_n = factor(lambda_n,
                           labels = c("lambda[n]: 500",
                                      "lambda[n]: 2000",
                                      "lambda[n]: 5000"))) %>%
  ggplot(aes(abundance, log(time), color = method, 
             group = interaction(abundance, method))) +
  geom_point() +
  theme_minimal() + 
  theme(legend.position = "top",
        legend.title = element_blank()) +
  geom_line(aes(group = method)) +
  labs(x = "abundance",
       y = "median log(seconds)") +
  facet_wrap(~lambda_n, labeller = label_parsed)
  
```




```{r fig_sim_expectation, fig.width=10, fig.height=5}
p_doc / p_times
```



