---
title: "simulation results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the paper. Libraries and code are sourced below. 

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(spatstat.random)
library(patchwork)
library(ggbreak)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```


# Expectation

## Clean and process simulation data

Here we load and process simulation results. Full simulations involve 36 scenarios with 50 simulated datasets each.


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "univariate_expectation", 
                              "varyAbundance_scSim", "20241007"), 
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)



```




```{r}
# make comp times dataset
times_df = df %>%
  select(r, iter, type, lambda_n,
         abundance,
         method,
         time) %>%
  filter(r == 0) %>%
  mutate(method = factor(method,
                         levels = c("k", "kinhom", "perm", 
                                    "kamp",  
                                    "kamp_lite"),
                         labels = c("K", "Kinhom","perm1000", "KAMP",
                                    "KAMP lite")),
         lambda_m = abundance * lambda_n) %>%
  group_by(method, type, lambda_n, lambda_m) %>%
  summarize(time = median(time)) %>%
  ungroup() 

# degree of clustering
doc_df = df %>%
  select(r, iter, type, lambda_n,
         abundance,
         method, csr, trans) %>%
  filter(round(r,2) == 1.25) %>%
  group_by(iter, type, lambda_n, abundance, method) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(value = trans-csr) %>%
  mutate(method = factor(method,
                         levels = c("k", "kinhom", "perm", 
                                    "kamp",  
                                    "kamp_lite"),
                         labels = c("K", "Kinhom","perm1000", "KAMP",
                                    "KAMP lite")),
         type = factor(type, levels = c("hom", "inhom", "homClust", "inhomClust"),
                            labels = c("homogeneous",
                                       "inhomogeneous",
                                       "homogeneous + clustered",
                                       "inhomogeneous + clustered")))
```





## Degree of clustering


```{r}
doc_df %>%
  filter(lambda_n == 5000) %>%
  mutate(abundance = factor(abundance))  %>%
  ggplot(aes(abundance, value, group = interaction(abundance, method), fill = method)) +
  geom_boxplot() +
  labs(x = "abundance",
       y = "degree of clustering") +
  theme_minimal() +
  theme(legend.position = "none",
        legend.title = element_blank()) +
  facet_wrap(~type, scales = "free", nrow = 1)
```






## Computation time


```{r}
p_times = times_df %>%
  filter(type == "hom") %>%
  mutate(lambda_m = factor(lambda_m),
         lambda_n = factor(lambda_n,
                           labels = c("lambda[n]: 1000",
                                      "lambda[n]: 2000",
                                      "lambda[n]: 5000",
                                      "lambda[n]: 10000"))) %>%
  ggplot(aes(lambda_m, log(time), color = method, 
             group = interaction(lambda_m, method))) +
  geom_point() +
  theme_minimal() + 
  theme(legend.position = "top",
        legend.title = element_blank()) +
  geom_line(aes(group = method)) +
  labs(x = TeX("$lambda_m$"),
       y = "median log(seconds)") +
  facet_wrap(~lambda_n, labeller = label_parsed, scales= "free")
  
```




```{r fig_sim_expectation, fig.width=10, fig.height=5}
p_doc / p_times
```



