---
title: "Analysis of VectraPolarisData"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---


```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(patchwork)
library(broom)
library(viridis)
library(RESI)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Load data

Using ovarian cancer data from [VectraPolarisData](https://www.bioconductor.org/packages/release/data/experiment/html/VectraPolarisData.html) package. Data cleaning process shown in `source/vectraPolarisdata.R`.

```{r}
load(here::here("data", "processed_ovarian_data.Rda"))
```


# Process data analysis results


Source file for consolidating analyses

```{r}
source(here::here("source", "utils.R"))
```


```{r}
files = list.files(here::here("output", "vpData"))

# call merge_files_vectraPolaris() defined in utils.R
df = map_dfr(files, merge_files_vectraPolaris)

# make comp times dataset
times_df = df %>%
  select(sample_id, time_k, time_kepd, time_kepdThin, time_kperm) %>%
  distinct() %>%
  pivot_longer(time_k:time_kperm, 
               names_to = "method", values_to = "time",
               names_prefix = "time_") %>%
  mutate(method = factor(method, levels = c("k","kperm", "kepd", "kepdThin"),
                         labels = c("k", "perm", "epd", "epd thin")))
  

# make degree of clustering dataset for use in survival analysis
doc_df = df %>%
  filter(r > 5) %>%
  select(sample_id, r, ktheo, khat, kepd, kepdThin, kperm) %>%
  #filter(round(r) %in% c(50, 100, 125, 150)) %>%
  mutate(k = khat - ktheo,
         perm = khat - kperm,
         epd = khat - kepd,
         epdThin = khat - kepdThin,
          r = round(r)) %>%
  select(sample_id, r, k, epd) 



# get pvalues 
dfp = map_dfr(files, merge_files_vectraPolaris, variance = TRUE)

#dfp = dfp %>%
 # filter(r > 49.5, r < 51)
```


* Look at curves with my correction and with other correction
* Look at speed of calculating expected value for mine and permutation
* Look at p-values

# Analysis

## Survival analysis

Do across radii. Computing RESI is computationally slow.

```{r}
# calculate abundance of immune cells
p_immune = ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m = sum(immune == "immune"),
            p_immune = m/n) %>%
  ungroup() 


# use K but also control for age, stage, and abundance
clinical = patient_level_ovarian %>%
  select(sample_id,  age = age_at_diagnosis, stage = stage_bin,
         survival_time, event = death, primary) %>%
  filter(sample_id %in% unique(doc_df$sample_id)) %>%
  distinct() %>%
  left_join(p_immune) %>%
  left_join(doc_df) %>%
  filter(primary == 1)




get_coxph = function(dat, rval, epd = TRUE){
  dat = filter(dat, r == rval)
  
  if(epd){
      mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + epd, 
              data = dat)
  }else{
      mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + k, 
              data = dat)
  }
  
  resi_object = resi(mod, data = dat)
  
  
  tidy(mod, exp = TRUE) %>%
    mutate(r = rval, epd = epd, 
           resi = coef(resi_object)$RESI,
           lower = coef(resi_object)$"2.5%",
           upper = coef(resi_object)$"97.5%")
}




filename = here::here("output",
                      "VPdata_coxPH_results.RData")

if(file.exists(filename)){
	load(filename)
}else{
  cph = bind_rows(map_dfr(unique(clinical$r), get_coxph, dat = clinical), 
                map_dfr(unique(clinical$r), get_coxph, dat = clinical, FALSE))
  
  save(cph, file = filename)
}


```


```{r}
cph %>%
  mutate(term = ifelse(term %in% c("k", "epd"), "fundiff", term)) %>%
  filter(term == "fundiff") %>%
  ggplot(aes(r, log(p.value), group = epd, color = epd)) +
  geom_hline(yintercept = log(0.05), linetype = 2) +
  geom_point()

```



```{r}
cph %>%
  mutate(term = ifelse(term %in% c("k", "epd"), "fundiff", term)) %>%
  filter(term == "fundiff") %>%
  ggplot(aes(r, resi, group = epd, color = epd)) +
  geom_point()
```





```{r}
clinical %>%
  as_tibble() %>%
  filter(r == 5) %>%
  select(sample_id, n,m,p_immune) %>%
  pivot_longer(n:p_immune, names_to = "stat", values_to = "value") %>%
  group_by(stat) %>%
  summarize(median = median(value),
            q25 = quantile(value, .25),
            q75 = quantile(value, .75)) %>%
  ungroup()

```


# Computation times

Computation times for expectation. 

```{r}
times_df %>%
  group_by(method) %>%
  summarize(median = median(time),
            max_seconds = max(time),
            "total (minutes)" = sum(time)/60) %>%
  ungroup() %>%
  arrange(median) %>%
  rename("median (seconds)" = median) %>%
  knitr::kable(digits = 1)
```


```{r}
times_df %>%
  #mutate(time = time/60) %>%
  filter(method != "k") %>%
  ggplot(aes(method, log(time), color = method, 
             group = method)) +
  geom_boxplot() +
  theme_minimal() + 
  coord_flip() +
  theme(legend.position = "none",
        legend.title = element_blank()) +
  labs(x = expression(lambda[m]),
       y = "log seconds") 
  
```


# Summary statistics

## Pvalues

```{r}
dfp %>%
  filter(sample_id %in% clinical$sample_id, r > 1) %>%
  select(sample_id, r, method, pvalue) %>% distinct() %>%
  group_by(r, method) %>%
  summarize(percent_sig_clustering = sum(pvalue<0.05)/n()) %>%
  ungroup() %>%
  filter(!is.na(percent_sig_clustering)) %>%
  ggplot(aes(r, percent_sig_clustering, group = method, color = method)) +
  geom_line()
```


# Figures


## Figure 1

This uses example data from the VectraPolarisData package. First data is loaded (contains three images).
Three ovarian images with K function and CSR.

Create code for plotting


```{r}
plot_k = function(idnum){
  dat = ovarian %>%
  filter(sample_id == idnum)

  dat_tumor = ovarian %>%
    filter(sample_id == idnum, tissue_category == "Tumor")
   
  p1 = dat %>%
    ggplot(aes(x, y)) +
    geom_point(size = 0.3, alpha = .5, color = "gray") +
    geom_point(data = filter(dat, immune == "immune"), 
               size = 0.3, color = "darkred", alpha = .5) +
    geom_point(data = dat_tumor, size = 0.5, alpha = .8, color = "gray") +
    geom_point(data = filter(dat_tumor, immune == "immune"), 
               size = 0.5, color = "darkred") +
    labs(x = "x position", y = "y position") +
    theme(axis.text = element_blank(), legend.position = "none") 
  

  p2 =   df %>%
    select(sample_id, r, k = khat, theo = ktheo, epd = kepd) %>%
    pivot_longer(k:epd, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, K != "epd", r <=150) %>%
    mutate(K = factor(K, levels = c("k", "theo"))) %>%
    ggplot(aes(r, value, group = K, color = K)) +
    geom_line() +
    scale_colour_manual(values = c("darkred", "blue"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$CSR_{\\pi r^2}$"))) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.25,.7),
          legend.title=element_blank())

  
  
  p3 = df %>%
    select(sample_id, r, k = khat, theo = ktheo, epd = kepd) %>%
    pivot_longer(k:epd, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, r <=150) %>%
    mutate(K = factor(K, levels = c("k", "epd", "theo"))) %>%
    ggplot(aes(r, value, group = K, color = K)) +
    geom_line() +
    scale_colour_manual(values = c("darkred", "purple", "blue"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$CSR_{epd}$"),
                                   theo = TeX("$CSR_{\\pi r^2}$"))) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.25,.7),
          legend.title=element_blank())

  
  list(p1, p2, p3)
  
}
```



Generate plots: before expected value statistic



```{r fig_3subjs, fig.height = 6}
high_k = "030120 P9HuP6 TMA 1-B_Core[1,7,E]_[14474,48365].im3"
low_k = "030120 P9HuP6 TMA 1-A_Core[1,6,C]_[10353,49826].im3"
patchy_k = "030120 P9HuP6 TMA 1-A_Core[1,1,C]_[10162,38841].im3"


high = plot_k(high_k)
low = plot_k(low_k)
patchy = plot_k(patchy_k)


(high[[1]] + low[[1]] + patchy[[1]]) / (high[[2]] + low[[2]] + patchy[[2]])

```




## Figure 5


Same as Figure 1, but including EPD CSR value

```{r fig_3subjs_epd, fig.height = 6}


(high[[1]] + low[[1]] + patchy[[1]]) / (high[[3]] + low[[3]] + patchy[[3]])

```



