---
title: "Analysis of VectraPolarisData"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---


```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(pec)
library(patchwork)
library(broom)
library(viridis)
library(RESI)
library(ggstance)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))


# Define a custom color palette for the methods
my_colors <- c(
  "K" = "#440154",
  "Kinhom" = "#414487",
  "perm" = "#2A788E",
  "KAMP" = "#7AD151",
  "KAMP lite" = "#FDE725"
)

```

# Load data

Using ovarian cancer data from [VectraPolarisData](https://www.bioconductor.org/packages/release/data/experiment/html/VectraPolarisData.html) package. Data cleaning process shown in `source/vectraPolarisdata.R`.

```{r}
load(here::here("data", "processed_ovarian_data.Rda"))
```


# Load bivariate data

```{r}
files = list.files(here::here("output", "vpData_bivariate"))

# call merge_files_vectraPolaris() defined in utils.R
df_biv = map_dfr(files, merge_files_vectraPolaris, bivariate = TRUE)

# make degree of clustering dataset for use in survival analysis
doc_biv = df_biv %>%
  filter(r >= 5) %>%
  select(sample_id, method, r, csr, trans) %>%
  mutate(doc = trans - csr) 


# get pvalues 
dfp = map_dfr(files, merge_files_vectraPolaris, variance = TRUE) %>%
  filter(method != "kperm approx") %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z),
         method = ifelse(method == "kperm", "perm", method),
         method = ifelse(method == "kamplite", "kamp_lite", method))  %>%
  select(sample_id, r, method, var, pvalue)

# get pvalues 
dfp_biv = map_dfr(files, merge_files_vectraPolaris, variance = TRUE, 
              bivariate = TRUE) %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z))  %>%
  select(-z, -Z)



dfp = left_join(doc_df, dfp) %>%
      mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")))
```


# Process data analysis results


Source file for consolidating analyses

```{r}
source(here::here("source", "utils.R"))
```


```{r}
files = list.files(here::here("output", "vpData"))

# call merge_files_vectraPolaris() defined in utils.R
df = map_dfr(files, merge_files_vectraPolaris)

# make comp times dataset
times_df = df %>%
  select(sample_id, method, time) %>%
  distinct() %>%
  mutate(method = factor(method, levels = c("k", "kinhom", "perm", "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")))
  

# make degree of clustering dataset for use in survival analysis
doc_df = df %>%
  filter(r >= 5) %>%
  select(sample_id, method, r, csr, trans) %>%
  mutate(doc = trans - csr) 


# get pvalues 
dfp = map_dfr(files, merge_files_vectraPolaris, variance = TRUE) %>%
  filter(method != "kperm approx") %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z),
         method = ifelse(method == "kperm", "perm", method),
         method = ifelse(method == "kamplite", "kamp_lite", method))  %>%
  select(sample_id, r, method, var, pvalue)


dfp = left_join(doc_df, dfp) %>%
      mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")))

```


* Look at curves with my correction and with other correction
* Look at speed of calculating expected value for mine and permutation
* Look at p-values

# Analysis




```{r}
# calculate abundance of immune cells
p_immune = ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m = sum(immune == "immune"),
            p_immune = m/n) %>%
  ungroup() 


# use K but also control for age, stage, and abundance
clinical = patient_level_ovarian %>%
  select(sample_id,  age = age_at_diagnosis, stage = stage_bin,
         survival_time, event = death, primary) %>%
  filter(sample_id %in% unique(doc_df$sample_id)) %>%
  distinct() %>%
  left_join(p_immune) %>%
  left_join(doc_df) %>%
  filter(primary == 1)


```





```{r}
clinical %>%
  as_tibble() %>%
  filter(r == 5) %>%
  select(sample_id, n,m,p_immune) %>%
  pivot_longer(n:p_immune, names_to = "stat", values_to = "value") %>%
  group_by(stat) %>%
  summarize(median = median(value),
            q25 = quantile(value, .25),
            q75 = quantile(value, .75),
            min = min(value),
            max = max(value)) %>%
  ungroup() %>%
  select(stat, min, q25, median, q75, max)

```



## Empirical densities


Here we generate empirical densities for use in data-driven simulation

```{r, eval = FALSE}
dat = filter(clinical, r == 101) %>%
  mutate(mean_k = mean((epd + k)/2),
         sd_k = sd((epd + k)/2),
         kamp = (epd - mean_k)/sd_k,
         k = (k - mean_k)/sd_k,
         kerror = k - kamp) %>%
  filter(kerror >= 0) %>%
  arrange(kerror) %>%
  select(sample_id, kerror, kamp, k)

# get combined std and mean, then calculate error and kamp distributions
# hist(dat$kerror)
# hist(dat$kamp)
# hist(dat$k)
# mean(dat$kamp)

# Estimate kernel density
density_kamp <- density(dat$kamp)
density_error = density(dat$kerror)

save(density_kamp, density_error, file = here::here("data", "densities_ovarian_data.Rda"))
```



# Computation times

Computation times for expectation. 

```{r}
times_df %>%
  group_by(method) %>%
  summarize(mean = mean(time),
            median = median(time),
            max_seconds = max(time),
            "total (minutes)" = sum(time)/60) %>%
  ungroup() %>%
  arrange(mean) %>%
  rename("mean (seconds)" = mean) %>%
  knitr::kable(digits = 1)
```



# Figures


## Figure 1

This uses example data from the VectraPolarisData package. First data is loaded (contains three images).
Three ovarian images with K function and CSR.

Create code for plotting


```{r}
plot_k = function(idnum){
  dat = ovarian %>%
  filter(sample_id == idnum)

  dat_tumor = ovarian %>%
    filter(sample_id == idnum#, tissue_category == "Tumor"
           )
   
  p1 = dat %>%
    ggplot(aes(x, y)) +
    geom_point(size = 0.3, alpha = .5, color = "gray") +
    geom_point(data = filter(dat, immune == "immune"), 
               size = 0.3, color = "darkred", alpha = .5) +
    geom_point(data = dat_tumor, size = 0.5, alpha = .8, color = "gray") +
    geom_point(data = filter(dat_tumor, immune == "immune"), 
               size = 0.5, color = "darkred") +
    labs(x = "x position", y = "y position") +
    theme(axis.text = element_blank(), legend.position = "none") 
  

 p2 =   df %>%
    filter(sample_id == idnum, method %in% c("k"), r <=150) %>%
    select(sample_id, r, k = trans, theo = csr) %>%
   pivot_longer(k:theo, names_to = "K", values_to = "value") %>%
    mutate(K = factor(K, levels = c("k", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.25) +
    scale_colour_manual(values = c("#440154", "gray"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$pi r^2$"))) +
    scale_linetype_manual(values = c(1, 2),
                          labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$\\pi r^2$"))) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.2,.75),
          legend.title=element_blank(),
          axis.text.y = element_blank(),
          legend.text = element_text(size = 12))

  
  
  p3 = df %>%
    select(-time) %>%
    filter(method %in% c("k", "kamp", "kamp_lite")) %>%
    pivot_wider(names_from = method, values_from = csr) %>%
    select(sample_id, r, theo = k, k = trans, epd = kamp, 
           epdlite = kamp_lite) %>%
   pivot_longer(theo:epdlite, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, r <=150) %>%
    mutate(K = factor(K, levels = c("k", "epd",  "epdlite", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.25) +
    scale_colour_manual(values = c("#440154", "#7AD151", "#FDE725", "gray"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$E_{KAMP}[\\hat{K}(r)]$"),
                                   epdlite = TeX("$E_{KAMPlite}[\\hat{K}(r)]$"),
                                   theo = TeX("$\\pi r^2$"))) +
        scale_linetype_manual(values = c(1, 4, 2, 5),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$E_{KAMP}[\\hat{K}(r)]$"),
                                   epdlite = TeX("$E_{KAMPlite}[\\hat{K}(r)]$"),
                                   theo = TeX("$\\pi r^2$"))
                        ) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.25,.7),
          legend.title=element_blank(),
          axis.text.y = element_blank(),
          legend.text = element_text(size = 12),
          legend.background = element_rect(fill = "transparent", color = NA))


  
  list(p1, p2, p3)
  
}
```



Generate plots: before expected value statistic



```{r fig_3subjs, fig.height = 6, fig.width = 12}
high_k = "030120 P9HuP6 TMA 1-B_Core[1,7,E]_[14474,48365].im3"
low_k = "030120 P9HuP6 TMA 1-A_Core[1,6,C]_[10353,49826].im3"
patchy_k = "030120 P9HuP6 TMA 1-A_Core[1,1,C]_[10162,38841].im3"
patchy_k2 = "030120 P9HuP6 TMA 1-B_Core[1,9,G]_[18855,52556].im3"

high = plot_k(high_k)
low = plot_k(low_k)
patchy = plot_k(patchy_k)
patchy_noClust = plot_k(patchy_k2)


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[2]], low[[2]], patchy[[2]], patchy_noClust[[2]]), ncol = 4)

```



## Figure 6


Same as Figure 1, but including KAMP CSR value

```{r fig_3subjs_epd, fig.height = 6, fig.width = 12}


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[3]], low[[3]], patchy[[3]], patchy_noClust[[3]]), ncol = 4)


```





## Figure 7


Boxplots of expectation, variance, and p-value for KAMP, KAMP lite, perm.

ALSO ADD DEGREE OF COLOCALIZATION TO THIS FIGURE

```{r fig_ovarian_within}
df_pval = dfp %>% 
  filter(sample_id %in% unique(clinical$sample_id)) %>%
  mutate(r = round(r),
         pvalue = log(pvalue),
         DoC = doc/1000,
         ) %>%
  filter(r %in% c(25, 101, 150, 200)
         ) %>%
  select(sample_id, r, method, DoC, pvalue) %>%
  pivot_longer(DoC:pvalue, names_to = "stat", values_to = "value") %>%
  mutate(r = factor(r),
         stat = factor(stat, levels = c("DoC", "var", "pvalue"),
                       labels = c("Degree of clustering", "Variance", "log(p-value)"))) 

df_pval %>%
  filter(value < 3750, value > -2000) %>%
  ggplot(aes(r, value, group = interaction(r, method), fill = method)) +
  geom_boxplot() +
  facet_wrap(~stat, scales = "free") +
  scale_fill_manual(values = my_colors) +
  labs(x = "radius") %>%
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.direction = "horizontal")
```


