---
title: "Bivariate analysis of VectraPolarisData"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---


```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(pec)
library(patchwork)
library(broom)
library(viridis)
library(RESI)
library(ggstance)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))


# Define a custom color palette for the methods
my_colors <- c(
  "K" = "#440154",
 # "perm" = "#2A788E",
  # perm wald = 22A884,
  "KAMP" = "#7AD151"
  #"KAMP lite" = "#FDE725"
)
```

# Load data

Using ovarian cancer data from [VectraPolarisData](https://www.bioconductor.org/packages/release/data/experiment/html/VectraPolarisData.html) package. Data cleaning process shown in `source/vectraPolarisdata.R`.

```{r}
load(here::here("data", "processed_ovarian_data.Rda"))
```

For the bivariate analysis we subset to subjects with at least 2 b-cells and at least 3 macrophages in the tumor compartment. This leaves a total of 71 subjects.

```{r}
ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  mutate(phenotype = case_when(phenotype_cd19 == "CD19+" ~ "immune1",
                                 phenotype_cd68 == "CD68+" ~ "immune2",
                                 TRUE ~ "other")) %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m_cd19 = sum(phenotype == "immune1"),
            m_cd68 = sum(phenotype == "immune2")) %>%
  ungroup()  %>%
  # only 71 subjects have at least 2 B-cells and at least 3 macrophages
  filter(m_cd68 >= 2, m_cd19 >= 2) %>%
  pivot_longer(n:m_cd68, names_to = "stat", values_to = "value") %>%
  group_by(stat) %>%
  summarize(median = median(value),
            q25 = quantile(value, .25),
            q75 = quantile(value, .75),
            min = min(value),
            max = max(value)) %>%
  ungroup() %>%
  select(stat, min, q25, median, q75, max)


```





# Process data analysis results


Source file for consolidating analyses

```{r}
source(here::here("source", "utils.R"))
```


```{r}
files = list.files(here::here("output", "vpData_bivariate"))

# call merge_files_vectraPolaris() defined in utils.R
df = map_dfr(files, merge_files_vectraPolaris, bivariate = TRUE)

# make comp times dataset
times_df = df %>%
  select(sample_id, time, method) %>%
  distinct() %>%
  mutate(method = factor(method, levels = c("k","perm", "kamp", "kamp_lite"),
                         labels = c("K", "perm", "KAMP", "KAMP lite")))
  

# make degree of clustering dataset for use in survival analysis
doc_df = df %>%
  filter(r >= 5) %>%
  select(sample_id, method, r, csr, trans) %>%
  mutate(doc = trans - csr) 



# get pvalues 
dfp = map_dfr(files, merge_files_vectraPolaris, variance = TRUE, bivariate = TRUE) %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z))  %>%
  select(-z, -Z)



```


* Look at curves with my correction and with other correction
* Look at speed of calculating expected value for mine and permutation
* Look at p-values

# Analysis

## Survival analysis

Do across radii. 

```{r}
# calculate abundance of immune cells
p_immune = ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m = sum(immune == "immune"),
            p_immune = m/n) %>%
  ungroup() 


# use K but also control for age, stage, and abundance
clinical = patient_level_ovarian %>%
  select(sample_id,  age = age_at_diagnosis, stage = stage_bin,
         survival_time, event = death, primary) %>%
  filter(sample_id %in% unique(doc_df$sample_id)) %>%
  distinct() %>%
  left_join(p_immune) %>%
  left_join(doc_df) %>%
  filter(primary == 1)

get_coxph = function(dat, rval, doc_method = "kamp", get_resi = TRUE){
  dat = filter(dat, r == rval, method == doc_method)  %>%
    mutate(doc = doc / 1000)
  
  mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + doc, 
              data = dat)
  
  if(get_resi){
    resi_object = resi(mod, data = dat)
    
      tidy(mod, exp = TRUE) %>%
    mutate(r = rval, method = doc_method, 
           resi = coef(resi_object)$RESI,
           lower = coef(resi_object)$"2.5%",
           upper = coef(resi_object)$"97.5%")
  }else{
      tidy(mod, exp = TRUE) %>%
      mutate(r = rval, method = doc_method)
  }
}

safe_get_coxph <- possibly(
  get_coxph,
  otherwise = tibble(term = NA, estimate = NA, std.error = NA, statistic = NA, p.value = NA, r = NA, method = NA,
                     resi = NA, lower = NA, upper = NA)
)




filename = here::here("output",
                      "VPdata_coxPH_results_bivariate.RData")

if(file.exists(filename)){
	load(filename)
}else{
  kamp = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp")
  k = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "k")
  kinhom = map_dfr(unique(clinical$r), safe_get_coxph, dat = clinical, doc_method = "kinhom")
  kamp_lite = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp_lite")
  perm = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "perm")
  
  cph = bind_rows(kamp, k, kinhom, kamp_lite, perm)
  
  kamp = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp", get_resi = FALSE)
  k = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "k", get_resi = FALSE)
  kinhom = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kinhom", get_resi = FALSE)
  kamp_lite = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp_lite", get_resi = FALSE)
  perm = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "perm", get_resi = FALSE)
  
  cph_noResi = bind_rows(kamp, k, kinhom, kamp_lite, perm)
  
  save(cph, cph_noResi, file = filename)
  
  
  ## save with no resi as well
}


```



# Table

Produce table of estimates at R = 75


```{r}
cph_noResi %>%
  filter(r == 100) %>%
  mutate(p = format.pval(p.value, digits = 2),
         lower = exp(log(estimate) - 1.96 * std.error),
         upper = exp(log(estimate) + 1.96 * std.error),
         OR = str_c(round(estimate, 2), " (", round(lower, 2), ", ", round(upper,2), ")")) %>%
  select(term, method, OR, p) %>%
  arrange(term) %>%
  knitr::kable(digits = 2)
```



Add to this plot: 

- edit labels
- figure out what is going on with Kinhom
- make colors consistent with previous plots 
- consider hazard ratio
- probably get rid of effect size stuff

```{r}
cph_noResi %>%
  filter(r == 100) %>%
  mutate(
    lower = exp(log(estimate) - 1.96 * std.error),
    upper = exp(log(estimate) + 1.96 * std.error),
    sig = case_when(
  p.value < 0.001 ~ "***",
  p.value < 0.01 ~ "**",
  p.value < 0.05 ~ "*",
  TRUE ~ "")
  ) %>%
  ggplot(aes(x = method, y = estimate, color = method)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_text(aes(label = sig), vjust = -1.2, color = "black", size = 6) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "grey40") +
  facet_wrap(~ term, scales = "free_y") +
  labs(
    x = "Method",
    y = "Hazard Ratio",
    title = "Hazard Ratios by Method for Each Term"
  ) +
  theme_minimal()  +
  theme(legend.position = "none") 
  
```


# Figures



## Figure 8




Think about how to expand this plot. Also want areas of significance.


```{r}
cph = cph %>% 
  mutate(significant = case_when(
      method == "k" & p.value < 0.05 ~ "K",
      method == "kamp" & p.value < 0.05 ~ "KAMP",
      TRUE ~ NA),
      significant = factor(significant, levels = c("K", "KAMP")),
      significant_es = case_when(
      method == "k" & lower > 0  ~ "K",
      method == "kamp"  & lower > 0 ~ "KAMP",
      TRUE ~ NA),
      method = factor(method, levels = c("k", "kamp"), labels = c("K", "KAMP")))
```



```{r}
es = cph %>%
  filter(term == "doc") %>%
  ggplot(aes(r, resi, group = method, color = method)) +
    geom_ribbon(
    data = filter(cph, term == "doc",significant_es == "K"),
    aes(ymin = 0, ymax = lower, fill = "K"),
    alpha = 0.2,
    color = NA
  ) +
    geom_ribbon(
    data = filter(cph, term == "doc",significant_es == "KAMP"),
    aes(ymin = 0, ymax = lower, fill = "KAMP"),
    alpha = 0.2,
    color = NA
  ) +
  geom_line() +
  geom_line(aes(y = upper), linetype = 3) +
  geom_line(aes(y = lower), linetype = 3) +
  scale_fill_manual(values = my_colors, guide = "none") +
   scale_color_manual(values = my_colors) +
  labs(x = "radius", y = "Effect size",
       title = "Standardized effect size for degree of colocalization") +
  theme(legend.title = element_blank(), legend.position = "none",
        legend.text = element_text(size = 12)) +
   geom_hline(yintercept = 0, linetype = 2, color = "gray")
```


```{r}
pp = cph %>%
  filter(term == "doc") %>%
  ggplot(aes(r, p.value, group = method, color = method)) +
  geom_ribbon(
    data = filter(cph, term == "doc",significant == "K"),
    aes(ymin = 0, ymax = p.value, fill = "K"),
    alpha = 0.2,
    color = NA
  ) +
    geom_ribbon(
    data = filter(cph, term == "doc",significant == "KAMP"),
    aes(ymin = 0, ymax = p.value, fill = "KAMP"),
    alpha = 0.2,
    color = NA
  ) +
  geom_line() +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors, guide = "none") +
  labs(x = "radius", y = "P-value",
       title = "P-value for degree of colocalization hazard ratio") +
  theme(legend.title = element_blank(), legend.position = c(0.6, 0.4),
        legend.text = element_text(size = 12)) +
  geom_hline(yintercept = 0.05, linetype = 2, color = "gray")
```



```{r fig_vpData_survival_biv, fig.height=6}
es/pp
```




```{r fig_vpData_survival, fig.width = 9}

p_df = cph %>%
  filter(r %in% c(25, 125, 175), term == "doc") %>%
  mutate(term = factor(term, label = "DoC")) %>%
  mutate(#p = format.pval(p.value, digits = 1),
         p = round(p.value, 2),
        resi = -0.1) %>%
  select(term, r, method, p,  resi)

cph %>%
  filter(r %in% c(25, 125, 175)) %>%
  select(term, r, resi, lower, upper, p.value, method) %>%
  mutate(term = factor(term,
                       levels = c("age", "stage", "p_immune", "doc"),
                       labels = c("age", "stage", "abundance", "DoC"))) %>%
  ggplot(aes(term, resi, color = method, 
             group = interaction(term, method), 
             linetype = method)) +
  geom_point(size = 2, position = position_dodge(0.5)) +
  theme_bw() +
  labs(y = "standardized effect size") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .7,
                position = position_dodge(0.5)) +
  theme(axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45,  hjust=1,
                                   size = 12),
        strip.text = element_text(size = 12),
        axis.title.y = element_text(size = 13),
        legend.title = element_blank(),
        legend.position = c(0.1, 0.15)) +
  geom_text(data = p_df, aes(label = p),  
            position = position_dodgev(0.1)) +
  geom_hline(yintercept = 0, linetype = 3) +
  scale_color_manual(values = my_colors) +
  facet_wrap(~r, labeller = label_both)
```

