---
title: "Analysis of VectraPolarisData"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---


```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(pec)
library(patchwork)
library(broom)
library(viridis)
library(RESI)
library(ggstance)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))


# Define a custom color palette for the methods
my_colors <- c(
  "K" = "#440154",
  "perm" = "#2A788E",
  # perm wald = 22A884,
  "KAMP" = "#7AD151",
  "KAMP lite" = "#FDE725"
)
```

# Load data

Using ovarian cancer data from [VectraPolarisData](https://www.bioconductor.org/packages/release/data/experiment/html/VectraPolarisData.html) package. Data cleaning process shown in `source/vectraPolarisdata.R`.

```{r}
load(here::here("data", "processed_ovarian_data.Rda"))
```


# Process data analysis results


Source file for consolidating analyses

```{r}
source(here::here("source", "utils.R"))
```


```{r}
files = list.files(here::here("output", "vpData"))

# call merge_files_vectraPolaris() defined in utils.R
df = map_dfr(files, merge_files_vectraPolaris)

# make comp times dataset
times_df = df %>%
  select(sample_id, time_k, time_kepd, time_kepdThin, time_kperm) %>%
  distinct() %>%
  pivot_longer(time_k:time_kperm, 
               names_to = "method", values_to = "time",
               names_prefix = "time_") %>%
  mutate(method = factor(method, levels = c("k","kperm", "kepd", "kepdThin"),
                         labels = c("K", "perm", "KAMP", "KAMP lite")))
  

# make degree of clustering dataset for use in survival analysis
doc_df = df %>%
  filter(r > 5) %>%
  select(sample_id, r, ktheo, khat, kepd, kepdThin, kperm) %>%
  #filter(round(r) %in% c(50, 100, 125, 150)) %>%
  mutate(k = khat - ktheo,
         perm = khat - kperm,
         epd = khat - kepd,
         epdThin = khat - kepdThin,
          r = round(r)) %>%
  select(sample_id, r, k, epd) 



# get pvalues 
dfp = map_dfr(files, merge_files_vectraPolaris, variance = TRUE) %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z))  %>%
  select(-z)

# calculate DOC for the variance data
dfp2 = dfp %>% 
  mutate(csr = pi * r^2,
         DoC = khat-expectation,
         k = khat - csr) %>%
  select(sample_id, r, method, DoC, k) %>%
  pivot_wider(names_from = method, values_from = DoC) %>%
  pivot_longer(k:kperm, names_to = "method", values_to = "DoC")
 

dfp = left_join(dfp2, dfp)
```


* Look at curves with my correction and with other correction
* Look at speed of calculating expected value for mine and permutation
* Look at p-values

# Analysis

## Survival analysis

Do across radii. 

```{r}
# calculate abundance of immune cells
p_immune = ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m = sum(immune == "immune"),
            p_immune = m/n) %>%
  ungroup() 


# use K but also control for age, stage, and abundance
clinical = patient_level_ovarian %>%
  select(sample_id,  age = age_at_diagnosis, stage = stage_bin,
         survival_time, event = death, primary) %>%
  filter(sample_id %in% unique(doc_df$sample_id)) %>%
  distinct() %>%
  left_join(p_immune) %>%
  left_join(doc_df) %>%
  filter(primary == 1)




get_coxph = function(dat, rval, epd = TRUE){
  dat = filter(dat, r == rval)
  
  if(epd){
      mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + epd, 
              data = dat)
  }else{
      mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + k, 
              data = dat)
  }
  
  resi_object = resi(mod, data = dat)
  
  
  tidy(mod, exp = TRUE) %>%
    mutate(r = rval, epd = epd, 
           resi = coef(resi_object)$RESI,
           lower = coef(resi_object)$"2.5%",
           upper = coef(resi_object)$"97.5%")
}




filename = here::here("output",
                      "VPdata_coxPH_results.RData")

if(file.exists(filename)){
	load(filename)
}else{
  cph = bind_rows(map_dfr(unique(clinical$r), get_coxph, dat = clinical), 
                map_dfr(unique(clinical$r), get_coxph, dat = clinical, FALSE))
  
  save(cph, file = filename)
}


```





```{r}
clinical %>%
  as_tibble() %>%
  filter(r == 5) %>%
  select(sample_id, n,m,p_immune) %>%
  pivot_longer(n:p_immune, names_to = "stat", values_to = "value") %>%
  group_by(stat) %>%
  summarize(median = median(value),
            q25 = quantile(value, .25),
            q75 = quantile(value, .75),
            min = min(value),
            max = max(value)) %>%
  ungroup() %>%
  select(stat, min, q25, median, q75, max)

```


## Out of sample survival prediction

Here we do cross validation to get out of sample predictions using the model with K and KAMP, and assess predictive performance.  Choosing 3 radii.

```{r, eval = FALSE}
dat = filter(clinical, r == 50) %>%
  mutate(kerror = k - epd) %>%
  filter(kerror >=0 ) %>%
  mutate(kamp_std = (epd -mean(epd))/sd(epd),
         kerror = log(kerror),
         kerror_std = (kerror - mean(kerror))/sd(kerror))
  
hist(dat$kerror)
hist(dat$epd)

var(log(dat$epd), na.rm = TRUE)
var(log(dat$k), na.rm = TRUE)

mean(dat$kerror)
var(dat$kerror)
cor(dat$kamp_std, dat$p_immune). # correlation is 0.3


mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + log(epd), 
              data = dat, x = TRUE, y = TRUE)

mod_k = coxph(Surv(survival_time, event) ~ age + stage + p_immune + k, 
              data = filter(clinical, r == 50), x = TRUE, y = TRUE)




# Compute cross-validated Brier scores
# Use pec to create prediction error curves with cross-validation
brier_scores <- pec(
  object = list("KAMP" = mod, "K" = mod_k),
  formula = Surv(survival_time, event) ~ age + stage + p_immune + epd,
  data =  filter(clinical, r == 50),
  times = seq(20, 120, by = 10), # Specify time points for evaluation
  exact = TRUE,
  splitMethod = "cv10"
)

plot(brier_scores, xlab = "Time", ylab = "Brier Score", main = "Cross-Validated Brier Scores")
```


# Computation times

Computation times for expectation. 

```{r}
times_df %>%
  group_by(method) %>%
  summarize(mean = mean(time),
            median = median(time),
            max_seconds = max(time),
            "total (minutes)" = sum(time)/60) %>%
  ungroup() %>%
  arrange(mean) %>%
  rename("mean (seconds)" = mean) %>%
  knitr::kable(digits = 1)
```


```{r}
times_df %>%
  #mutate(time = time/60) %>%
  filter(method != "k") %>%
  ggplot(aes(method, log(time), color = method, 
             group = method)) +
  geom_boxplot() +
  theme_minimal() + 
  coord_flip() +
  theme(legend.position = "none",
        legend.title = element_blank()) +
  labs(x = expression(lambda[m]),
       y = "log seconds") 
  
```





# Figures


## Figure 1

This uses example data from the VectraPolarisData package. First data is loaded (contains three images).
Three ovarian images with K function and CSR.

Create code for plotting


```{r}
plot_k = function(idnum){
  dat = ovarian %>%
  filter(sample_id == idnum)

  dat_tumor = ovarian %>%
    filter(sample_id == idnum, tissue_category == "Tumor")
   
  p1 = dat %>%
    ggplot(aes(x, y)) +
    geom_point(size = 0.3, alpha = .5, color = "gray") +
    geom_point(data = filter(dat, immune == "immune"), 
               size = 0.3, color = "darkred", alpha = .5) +
    geom_point(data = dat_tumor, size = 0.5, alpha = .8, color = "gray") +
    geom_point(data = filter(dat_tumor, immune == "immune"), 
               size = 0.5, color = "darkred") +
    labs(x = "x position", y = "y position") +
    theme(axis.text = element_blank(), legend.position = "none") 
  

 p2 =   df %>%
    select(sample_id, r, k = khat, theo = ktheo, epd = kepd) %>%
    pivot_longer(k:epd, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, K != "epd", r <=150) %>%
    mutate(K = factor(K, levels = c("k", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.25) +
    scale_colour_manual(values = c("darkred", "#440154"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$pi r^2$"))) +
    scale_linetype_manual(values = c(1, 2),
                          labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$\\pi r^2$"))) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.2,.75),
          legend.title=element_blank(),
          axis.text.y = element_blank(),
          legend.text = element_text(size = 12))

  
  
  p3 = df %>%
    select(sample_id, r, k = khat, theo = ktheo, epd = kepd) %>%
    pivot_longer(k:epd, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, r <=150) %>%
    mutate(K = factor(K, levels = c("k", "epd", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.25) +
    scale_colour_manual(values = c("darkred", "#7AD151", "#440154"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$KAMP$"),
                                   theo = TeX("$\\pi r^2$"))) +
        scale_linetype_manual(values = c(1, 4, 2),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$KAMP$"),
                                   theo = TeX("$\\pi r^2$"))) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.25,.7),
          legend.title=element_blank(),
          axis.text.y = element_blank(),
          legend.text = element_text(size = 12),
          legend.background = element_rect(fill = "transparent", color = NA))


  
  list(p1, p2, p3)
  
}
```



Generate plots: before expected value statistic



```{r fig_3subjs, fig.height = 6, fig.width = 12}
high_k = "030120 P9HuP6 TMA 1-B_Core[1,7,E]_[14474,48365].im3"
low_k = "030120 P9HuP6 TMA 1-A_Core[1,6,C]_[10353,49826].im3"
patchy_k = "030120 P9HuP6 TMA 1-A_Core[1,1,C]_[10162,38841].im3"
patchy_k2 = "030120 P9HuP6 TMA 1-B_Core[1,9,G]_[18855,52556].im3"

high = plot_k(high_k)
low = plot_k(low_k)
patchy = plot_k(patchy_k)
patchy_noClust = plot_k(patchy_k2)


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[2]], low[[2]], patchy[[2]], patchy_noClust[[2]]), ncol = 4)

```


```{r, eval = FALSE}
library(plotly)
p = dfp %>%
  filter(r <=150, method %in% c("k", "kepd"),
         round(r) %in% c( 125)) %>%
  select(sample_id, r, method, DoC) %>%
  pivot_wider(names_from = method, values_from = DoC) %>%
  # sound k or divide by 1000 to make this more manageable 
  mutate(k = round(k/1000),
         kepd = round(kepd/1000)) %>%
  mutate(diff = k - kepd) %>%
  filter(diff > 0) %>%
  ggplot(aes(kepd, diff, color = sample_id)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point()


ggplotly(p)


ids = dfp %>%
  filter(r <=150, method %in% c("k", "kepd"),
         round(r) %in% c( 125)) %>%
  select(sample_id, r, method, DoC) %>%
  pivot_wider(names_from = method, values_from = DoC) %>%
  # sound k or divide by 1000 to make this more manageable 
  mutate(k = round(k/1000),
         kepd = round(kepd/1000))%>%
  filter(kepd %in% c(16, -43, -42, 2, 24, 8, 7, 3, 4)) %>%
  pull(sample_id)

patchy2 = plot_k(ids[26])

# 26!

patchy2[[1]] / patchy2[[3]]
```




## Figure 4


Same as Figure 1, but including EPD CSR value

```{r fig_3subjs_epd, fig.height = 6, fig.width = 12}


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[3]], low[[3]], patchy[[3]], patchy_noClust[[3]]), ncol = 4)


```





## Figure 5

Boxplots of expectation, variance, and p-value for KAMP, KAMP lite, perm

```{r fig_ovarian_within}
dfp %>% 
  mutate(r = round(r),
         var = log(var),
         pvalue = log(pvalue)) %>%
  filter(r %in% c(25, 125, 175)
         #r == 125
         ) %>%
  select(sample_id, r, method, DoC, var, pvalue) %>%
  pivot_longer(DoC:pvalue, names_to = "stat", values_to = "value") %>%
  #filter(!is.na(value)) %>%
  mutate(r = factor(r),
         stat = factor(stat, levels = c("DoC", "var", "pvalue"),
                       labels = c("Degree of clustering", "log(Variance)", "log(p-value)")),
         method = factor(method, levels = c("k", "kperm", "kepd", "kepdThin"),
                         labels = c("K", "perm", "KAMP", "KAMP lite"))) %>%
  filter(value < 400000) %>%
  ggplot(aes(r, value, group = interaction(r, method), fill = method)) +
  geom_boxplot() +
  facet_wrap(~stat, scales = "free") +
  scale_fill_manual(values = my_colors) +
  theme(legend.title = element_blank(),
        legend.position = c(0.1, 0.9),
        legend.direction = "horizontal")
```



## Figure 6



```{r}

cph = cph %>% 
  mutate(method = ifelse(epd == TRUE, "KAMP", "K"),
         term = ifelse(term %in% c("k", "epd"), "doc", term),
         r = round(r))
```




Think about how to expand this plot. Also want areas of significance.

```{r}
cph %>%
  filter(term == "doc") %>%
  ggplot(aes(r, resi, group = method, color = method)) +
  geom_line() +
  geom_line(aes(y = upper), linetype = 2) +
  geom_line(aes(y = lower), linetype = 2)
```






```{r fig_vpData_survival, fig.width = 9}

p_df = cph %>%
  filter(r %in% c(25, 125, 175), term == "doc") %>%
  mutate(term = factor(term, label = "DoC")) %>%
  mutate(#p = format.pval(p.value, digits = 1),
         p = round(p.value, 2),
        resi = -0.1) %>%
  select(term, r, method, p,  resi)

cph %>%
  filter(r %in% c(25, 125, 175)) %>%
  select(term, r, resi, lower, upper, p.value, method) %>%
  mutate(term = factor(term,
                       levels = c("age", "stage", "p_immune", "doc"),
                       labels = c("age", "stage", "abundance", "DoC"))) %>%
  ggplot(aes(term, resi, color = method, 
             group = interaction(term, method), 
             linetype = method)) +
  geom_point(size = 2, position = position_dodge(0.5)) +
  theme_bw() +
  labs(y = "standardized effect size") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .7,
                position = position_dodge(0.5)) +
  theme(axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45,  hjust=1,
                                   size = 12),
        strip.text = element_text(size = 12),
        axis.title.y = element_text(size = 13),
        legend.title = element_blank(),
        legend.position = c(0.1, 0.15)) +
  geom_text(data = p_df, aes(label = p),  
            position = position_dodgev(0.1)) +
  geom_hline(yintercept = 0, linetype = 3) +
  scale_color_manual(values = my_colors) +
  facet_wrap(~r, labeller = label_both)
```

