---
title: "Analysis of VectraPolarisData"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---



```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(patchwork)
library(broom)
library(viridis)
library(RESI)
library(ggstance)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))


# Define a custom color palette for the methods

my_colors <- c(
  "K"         = "#6A3D9A",  # purple
  "Kinhom"    = "#1F78B4",  # blue
  "perm"      = "#009E73",  # bluish green
  "KAMP"      = "#D55E00",  # vermillion
  "KAMP lite" = "#E69F00"   # orange
)

```

# Load univariate data

Using ovarian cancer data from [VectraPolarisData](https://www.bioconductor.org/packages/release/data/experiment/html/VectraPolarisData.html) package. Data cleaning process shown in `source/vectraPolarisdata.R`.

```{r}
load(here::here("data", "processed_ovarian_data.Rda"))
```


# Process data analysis results


Source file for consolidating analyses

```{r}
source(here::here("source", "utils.R"))
```


```{r}
files = list.files(here::here("output", "vpData"))

# call merge_files_vectraPolaris() defined in utils.R
df = map_dfr(files, merge_files_vectraPolaris)

# make comp times dataset
times_df = df %>%
  select(sample_id, method, time) %>%
  distinct() %>%
  mutate(method = factor(method, levels = c("k", "kinhom", "perm", "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")))
  

# make degree of clustering dataset for use in survival analysis
doc_df = df %>%
  filter(r >= 5) %>%
  select(sample_id, method, r, csr, trans) %>%
  mutate(doc = trans - csr) 


# get pvalues 
dfp = map_dfr(files, merge_files_vectraPolaris, variance = TRUE) %>%
  filter(method != "kperm approx") %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z),
         method = ifelse(method == "kperm", "perm", method),
         method = ifelse(method == "kamplite", "kamp_lite", method))  %>%
  select(sample_id, r, method, var, pvalue)


dfp = left_join(doc_df, dfp) %>%
      mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")))

```


# Load and process bivariate data


```{r}

ids_biv = ovarian %>%
    filter(tissue_category == "Tumor") %>%
  mutate(phenotype = case_when(phenotype_cd19 == "CD19+" ~ "immune1",
                                 phenotype_cd68 == "CD68+" ~ "immune2",
                                 TRUE ~ "other")) %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m_cd19 = sum(phenotype == "immune1"),
            m_cd68 = sum(phenotype == "immune2")) %>%
  ungroup()  %>%
  # only 71 subjects have at least 2 B-cells and at least 3 macrophages
  filter(m_cd68 >= 2, m_cd19 >= 2) %>% pull(sample_id)
  
  
files = list.files(here::here("output", "vpData_bivariate"))

# call merge_files_vectraPolaris() defined in utils.R
df_biv = map_dfr(files, merge_files_vectraPolaris, bivariate = TRUE)

# make degree of clustering dataset for use in survival analysis
doc_biv = df_biv %>%
  filter(r >= 5) %>%
  select(sample_id, method, r, csr, trans) %>%
  mutate(doc = trans - csr) 


# get pvalues 
dfp_biv = map_dfr(files, merge_files_vectraPolaris, variance = TRUE, 
              bivariate = TRUE) %>%
    mutate(z = (khat-expectation) / sqrt(var),
         pvalue = pnorm(-z),
                  method = ifelse(method == "kperm", "perm", method),
         method = ifelse(method == "kamplite", "kamp_lite", method))  %>%
  select(sample_id, r, method, var, pvalue)



dfp_biv = left_join(doc_biv, dfp_biv) %>%
      mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite"))) %>%
  filter(sample_id %in% ids_biv)
```


# Analysis




```{r}
# calculate abundance of immune cells
p_immune = ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m = sum(immune == "immune"),
            p_immune = m/n) %>%
  ungroup() 


# use K but also control for age, stage, and abundance
clinical = patient_level_ovarian %>%
  select(sample_id,  age = age_at_diagnosis, stage = stage_bin,
         survival_time, event = death, primary) %>%
  filter(sample_id %in% unique(doc_df$sample_id)) %>%
  distinct() %>%
  left_join(p_immune) %>%
  left_join(doc_df) %>%
  filter(primary == 1)


```





```{r}
clinical %>%
  as_tibble() %>%
  filter(r == 5) %>%
  select(sample_id, n,m,p_immune) %>%
  pivot_longer(n:p_immune, names_to = "stat", values_to = "value") %>%
  group_by(stat) %>%
  summarize(median = median(value),
            q25 = quantile(value, .25),
            q75 = quantile(value, .75),
            min = min(value),
            max = max(value)) %>%
  ungroup() %>%
  select(stat, min, q25, median, q75, max)

```





# Computation times

Computation times for expectation. 

```{r}
times_df %>%
  group_by(method) %>%
  summarize(mean = mean(time),
            median = median(time),
            max_seconds = max(time),
            "total (minutes)" = sum(time)/60) %>%
  ungroup() %>%
  arrange(mean) %>%
  rename("mean (seconds)" = mean) %>%
  knitr::kable(digits = 1)
```



# Figures


## Figure 1

This uses example data from the VectraPolarisData package. First data is loaded (contains three images).
Three ovarian images with K function and CSR.

Create code for plotting


```{r}
plot_k = function(idnum){
  dat = ovarian %>%
  filter(sample_id == idnum)

  dat_tumor = ovarian %>%
    filter(sample_id == idnum#, tissue_category == "Tumor"
           )
   
  p1 = dat %>%
    ggplot(aes(x, y)) +
    geom_point(size = 0.3, alpha = .5, color = "gray") +
    geom_point(data = filter(dat, immune == "immune"), 
               size = 0.3, color = "darkred", alpha = .5) +
    geom_point(data = dat_tumor, size = 0.5, alpha = .8, color = "gray") +
    geom_point(data = filter(dat_tumor, immune == "immune"), 
               size = 0.5, color = "darkred") +
    labs(x = "x position", y = "y position") +
    theme(axis.text = element_blank(), legend.position = "none",
          axis.title = element_text(size = 12)) 
  

 p2 =   df %>%
    filter(sample_id == idnum, method %in% c("k"), r <=150) %>%
    select(sample_id, r, k = trans, theo = csr) %>%
   pivot_longer(k:theo, names_to = "K", values_to = "value") %>%
    mutate(K = factor(K, levels = c("k", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.75) +
    scale_colour_manual(values = c("#440154", "gray"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$pi r^2$"))) +
    scale_linetype_manual(values = c(1, 2),
                          labels = c(k = TeX("$\\hat{K}(r)$"),
                                   theo = TeX("$\\pi r^2$"))) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.2,.75),
          legend.title=element_blank(),
          axis.title = element_text(size = 14),
          axis.text.y = element_blank(),
          legend.text = element_text(size = 14))

  
  
  p3 = df %>%
    select(-time) %>%
    filter(method %in% c("k", "kamp", "kamp_lite")) %>%
    pivot_wider(names_from = method, values_from = csr) %>%
    select(sample_id, r, theo = k, k = trans, epd = kamp, 
           epdlite = kamp_lite) %>%
   pivot_longer(theo:epdlite, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, r <=150) %>%
    mutate(K = factor(K, levels = c("k", "epd",  "epdlite", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.75) +
    scale_colour_manual(values = c("#6A3D9A", "#D55E00", "#E69F00", "gray"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$E_{KAMP}[\\hat{K}(r)]$"),
                                   epdlite = TeX("$E_{KAMPlite}[\\hat{K}(r)]$"),
                                   theo = TeX("$\\pi r^2$"))) +
        scale_linetype_manual(values = c(1, 4, 2, 5),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$E_{KAMP}[\\hat{K}(r)]$"),
                                   epdlite = TeX("$E_{KAMPlite}[\\hat{K}(r)]$"),
                                   theo = TeX("$\\pi r^2$"))
                        ) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.25,.7),
          legend.title=element_blank(),
          axis.title = element_text(size = 14),
          axis.text.y = element_blank(),
          legend.text = element_text(size = 14),
          legend.background = element_rect(fill = "transparent", color = NA))


    
  p4 = df %>%
    select(-time) %>%
    filter(method %in% c("k", "kamp", "kamp_lite")) %>%
    pivot_wider(names_from = method, values_from = csr) %>%
    select(sample_id, r, theo = k, k = trans, epd = kamp, 
           epdlite = kamp_lite) %>%
   pivot_longer(theo:epdlite, names_to = "K", values_to = "value") %>%
    filter(sample_id == idnum, r <=150, K != "epdlite") %>%
    mutate(K = factor(K, levels = c("k", "epd",  "epdlite", "theo"))) %>%
    ggplot(aes(r, value, color = K, linetype = K)) +
    geom_line(linewidth = 1.75) +
    scale_colour_manual(values = c("#6A3D9A", "#D55E00", "#E69F00", "gray"),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$E_{KAMP}[\\hat{K}(r)]$"),
                                   epdlite = TeX("$E_{KAMPlite}[\\hat{K}(r)]$"),
                                   theo = TeX("$\\pi r^2$"))) +
        scale_linetype_manual(values = c(1, 4, 2, 5),
                        labels = c(k = TeX("$\\hat{K}(r)$"),
                                   epd = TeX("$E_{KAMP}[\\hat{K}(r)]$"),
                                   epdlite = TeX("$E_{KAMPlite}[\\hat{K}(r)]$"),
                                   theo = TeX("$\\pi r^2$"))
                        ) +
    labs(y = TeX("K(r)")) +
    theme(legend.position = c(.25,.7),
          legend.title=element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_text(size = 14),
          legend.text = element_text(size = 14),
          legend.background = element_rect(fill = "transparent", color = NA))

  
  list(p1, p2, p3, p4)
  
}
```



Generate plots: before expected value statistic



```{r fig_3subjs, fig.height = 6, fig.width = 12}
high_k = "030120 P9HuP6 TMA 1-B_Core[1,7,E]_[14474,48365].im3"
low_k = "030120 P9HuP6 TMA 1-A_Core[1,6,C]_[10353,49826].im3"
patchy_k = "030120 P9HuP6 TMA 1-A_Core[1,1,C]_[10162,38841].im3"
patchy_k2 = "030120 P9HuP6 TMA 1-B_Core[1,9,G]_[18855,52556].im3"

high = plot_k(high_k)
low = plot_k(low_k)
patchy = plot_k(patchy_k)
patchy_noClust = plot_k(patchy_k2)


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[2]], low[[2]], patchy[[2]], patchy_noClust[[2]]), ncol = 4)

```



## Figure 6


Same as Figure 1, but including KAMP CSR value

```{r fig_3subjs_epd, fig.height = 6, fig.width = 12}


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[3]], low[[3]], patchy[[3]], patchy_noClust[[3]]), ncol = 4)


```


## Figure 6 talks


Pared down version for talk

```{r fig_3subjs_talk, fig.height = 6, fig.width = 12}


wrap_plots(list(high[[1]], low[[1]], patchy[[1]], patchy_noClust[[1]], high[[4]], low[[4]], patchy[[4]], patchy_noClust[[4]]), ncol = 4)


```



## Figure 7


Boxplots of expectation, variance, and p-value for KAMP, KAMP lite, perm.


```{r }
df_pval = dfp %>% 
  filter(sample_id %in% unique(clinical$sample_id)) %>%
  mutate(r = round(r),
         pvalue = log(pvalue),
         DoC = doc/1000,
         ) %>%
  filter(r %in% c(25, 50, 100, 150)
         ) %>%
  select(sample_id, r, method, DoC, pvalue) %>%
  pivot_longer(DoC:pvalue, names_to = "stat", values_to = "value") %>%
  mutate(r = factor(r),
         stat = factor(stat, levels = c("DoC", "var", "pvalue"),
                       labels = c("Degree of clustering", "Variance", "log(p-value)"))) 


df_pval_biv = dfp_biv %>% 
  filter(sample_id %in% unique(clinical$sample_id)) %>%
  mutate(r = round(r),
         pvalue = log(pvalue),
         DoC = doc/1000,
         ) %>%
  filter(r %in% c(25, 50, 100, 150)
         ) %>%
  select(sample_id, r, method, DoC, pvalue) %>%
  pivot_longer(DoC:pvalue, names_to = "stat", values_to = "value") %>%
  mutate(r = factor(r),
         stat = factor(stat, levels = c("DoC", "var", "pvalue"),
                       labels = c("Degree of clustering", "Variance", "log(p-value)"))) 

```


```{r fig_ovarian_within, fig.width = 10, fig.height = 6}
univ = df_pval %>%
  filter(value < 200, value > -2000) %>%
  ggplot(aes(r, value, group = interaction(r, method), fill = method)) +
  geom_boxplot() +
  facet_wrap(~stat, scales = "free") +
  scale_fill_manual(values = my_colors) +
  labs(x = "radius", title = "Degree of immune cell clustering") +
  theme(legend.position = "none")

biv = df_pval_biv %>%
  mutate(value = ifelse(value < -20 & stat == "log(p-value)",
                        -5, value)) %>%
  filter(value < 150, value > -99) %>%
  ggplot(aes(r, value, group = interaction(r, method), fill = method)) +
  geom_boxplot() +
  facet_wrap(~stat, scales = "free") +
  scale_fill_manual(values = my_colors) +
  labs(x = "radius", title = "Degree of B-cell and macrophage colocalization") +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.direction = "horizontal")

univ / biv
```



# Across Sample Analysis

## Bivariate cell counts

For the bivariate analysis we subset to subjects with at least 2 b-cells and at least 3 macrophages in the tumor compartment. This leaves a total of 71 subjects.

```{r}
ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  mutate(phenotype = case_when(phenotype_cd19 == "CD19+" ~ "immune1",
                                 phenotype_cd68 == "CD68+" ~ "immune2",
                                 TRUE ~ "other")) %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m_cd19 = sum(phenotype == "immune1"),
            m_cd68 = sum(phenotype == "immune2")) %>%
  ungroup()  %>%
  # only 71 subjects have at least 2 B-cells and at least 3 macrophages
  filter(m_cd68 >= 2, m_cd19 >= 2) %>%
  pivot_longer(n:m_cd68, names_to = "stat", values_to = "value") %>%
  group_by(stat) %>%
  summarize(median = median(value),
            q25 = quantile(value, .25),
            q75 = quantile(value, .75),
            min = min(value),
            max = max(value)) %>%
  ungroup() %>%
  select(stat, min, q25, median, q75, max)


```






## Survival analysis

```{r}
# calculate abundance of immune cells
p_immune = ovarian %>% 
  filter(tissue_category == "Tumor") %>%
  group_by(sample_id) %>%
  summarize(n = n(),
            m = sum(immune == "immune"),
            p_immune = m/n) %>%
  ungroup() 


# use K but also control for age, stage, and abundance
clinical = patient_level_ovarian %>%
  select(sample_id,  age = age_at_diagnosis, stage = stage_bin,
         survival_time, event = death, primary) %>%
  filter(sample_id %in% unique(doc_biv$sample_id)) %>%
  distinct() %>%
  left_join(p_immune) %>%
  left_join(doc_biv) %>%
  filter(primary == 1)

get_coxph = function(dat, rval, doc_method = "kamp", get_resi = TRUE){
  dat = filter(dat, r == rval, method == doc_method)  %>%
    mutate(doc = doc / 1000)
  
  mod = coxph(Surv(survival_time, event) ~ age + stage + p_immune + doc, 
              data = dat)
  
  if(get_resi){
    resi_object = resi(mod, data = dat)
    
      tidy(mod, exp = TRUE) %>%
    mutate(r = rval, method = doc_method, 
           resi = coef(resi_object)$RESI,
           lower = coef(resi_object)$"2.5%",
           upper = coef(resi_object)$"97.5%")
  }else{
      tidy(mod, exp = TRUE) %>%
      mutate(r = rval, method = doc_method, cindex = concordance(mod)$concordance)
  }
}

safe_get_coxph <- possibly(
  get_coxph,
  otherwise = tibble(term = NA, estimate = NA, std.error = NA, statistic = NA, p.value = NA, r = NA, method = NA,
                     resi = NA, lower = NA, upper = NA)
)




filename = here::here("output",
                      "VPdata_coxPH_results_bivariate.RData")

if(file.exists(filename)){
	load(filename)
}else{
  kamp = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp")
  k = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "k")
  kinhom = map_dfr(unique(clinical$r), safe_get_coxph, dat = clinical, doc_method = "kinhom")
  kamp_lite = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp_lite")
  perm = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "perm")
  
  cph = bind_rows(kamp, k, kinhom, kamp_lite, perm)
  
  kamp = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp", get_resi = FALSE)
  k = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "k", get_resi = FALSE)
  kinhom = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kinhom", get_resi = FALSE)
  kamp_lite = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "kamp_lite", get_resi = FALSE)
  perm = map_dfr(unique(clinical$r), get_coxph, dat = clinical, doc_method = "perm", get_resi = FALSE)
  
  cph_noResi = bind_rows(kamp, k, kinhom, kamp_lite, perm)
  
  save(cph, cph_noResi, file = filename)
  
  
}


```







# Figures



## Figure 8

```{r}
cindex = cph_noResi %>%
   mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")),
         term = factor(term, levels = c("doc", "p_immune", "age", "stage"))) %>%
  dplyr::select(r, method, cindex) %>%
  distinct() %>%
  ggplot(aes(r, cindex, group = method, color = method, linetype = method)) +
  geom_line(linewidth = 1.3) +
  labs(x = "Radius",y = "Concordance index")+
  theme(legend.position = c(0.7, 0.8),
        legend.title = element_blank(),
        legend.direction = "horizontal") +
  scale_color_manual(values = my_colors)  
```


```{r}
cph = cph %>% 
  filter(method %in% c("k", "kamp")) %>%
  mutate(significant = case_when(
      method == "k" & p.value < 0.05 ~ "K",
      method == "kamp" & p.value < 0.05 ~ "KAMP",
      TRUE ~ NA),
      significant = factor(significant, levels = c("K", "KAMP")),
      significant_es = case_when(
      method == "k" & lower > 0  ~ "K",
      method == "kamp"  & lower > 0 ~ "KAMP",
      TRUE ~ NA),
      method = factor(method, levels = c("k", "kamp"), labels = c("K", "KAMP")))
```



```{r}
r50 = cph_noResi %>%
  filter(r == 50, method != "kinhom"
         ) %>%
  mutate(
    lower = exp(log(estimate) - 1.96 * std.error),
    upper = exp(log(estimate) + 1.96 * std.error),
    sig = case_when(
  p.value < 0.001 ~ "***",
  p.value < 0.01 ~ "**",
  p.value < 0.05 ~ "*",
  TRUE ~ "")
  ) %>%
  mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")),
         term = factor(term, levels = c("doc", "p_immune", "age", "stage"),
                       labels = c("Degree of Colocalization at r = 50", "Proportion of immune cells",
                                  "Age", "Cancer stage"))) %>%
  ggplot(aes(x = method, y = estimate, color = method)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_text(aes(label = sig), vjust = -1.2, color = "black", size = 6) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "grey40") +
  facet_wrap(~ term, scales = "free_y", nrow = 1) +
  labs(
    x = "Method",
    y = "Hazard Ratio",
    title = "Hazard ratios for all model covariates at r = 50"
  ) +
  scale_color_manual(values = my_colors) +
  theme_minimal()  +
  theme(legend.position = "none") 
  
```



```{r}
es = cph %>%
  filter(term == "doc") %>%
  ggplot(aes(r, resi, group = method, color = method)) +
    geom_ribbon(
    data = filter(cph, term == "doc",significant_es == "K"),
    aes(ymin = 0, ymax = lower, fill = "K"),
    alpha = 0.2,
    color = NA
  ) +
    geom_ribbon(
    data = filter(cph, term == "doc",significant_es == "KAMP"),
    aes(ymin = 0, ymax = lower, fill = "KAMP"),
    alpha = 0.2,
    color = NA
  ) +
  geom_line(linewidth = 1.3) +
  geom_line(aes(y = upper), linetype = 3) +
  geom_line(aes(y = lower), linetype = 3) +
  scale_fill_manual(values = my_colors, guide = "none") +
   scale_color_manual(values = my_colors) +
  labs(x = "Radius", y = "Effect size",
       title = "Standardized effect size for degree of colocalization") +
  theme(legend.title = element_blank(), legend.position = "none",
        legend.text = element_text(size = 12)) +
   geom_hline(yintercept = 0, linetype = 2, color = "gray")
```


```{r}
pp = cph %>%
  filter(term == "doc") %>%
  ggplot(aes(r, p.value, group = method, color = method)) +
  geom_ribbon(
    data = filter(cph, term == "doc",significant == "K"),
    aes(ymin = 0, ymax = p.value, fill = "K"),
    alpha = 0.2,
    color = NA
  ) +
    geom_ribbon(
    data = filter(cph, term == "doc",significant == "KAMP"),
    aes(ymin = 0, ymax = p.value, fill = "KAMP"),
    alpha = 0.2,
    color = NA
  ) +
  geom_line(linewidth = 1.3) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors, guide = "none") +
  labs(x = "Radius", y = "P-value",
       title = "P-value for degree of colocalization hazard ratio") +
  theme(legend.title = element_blank(), legend.position = c(0.6, 0.4),
        legend.text = element_text(size = 12)) +
    geom_vline(xintercept = 50, linetype = 2, color = "gray") +
  geom_hline(yintercept = 0.05, linetype = 2, color = "gray")
```



```{r fig_vpData_survival_biv, fig.width = 10, fig.height=6}
pp/cindex/r50 + plot_layout(heights = c(1.5, 1.5, 1))
```


## Figure 8 talk


```{r}
r50 = cph_noResi %>%
  filter(r == 50, method != "kinhom"
         ) %>%
  mutate(
    lower = exp(log(estimate) - 1.96 * std.error),
    upper = exp(log(estimate) + 1.96 * std.error),
    sig = case_when(
  p.value < 0.001 ~ "***",
  p.value < 0.01 ~ "**",
  p.value < 0.05 ~ "*",
  TRUE ~ "")
  ) %>%
  mutate(method = factor(method, levels = c("k", "kinhom", "perm", 
                                                "kamp", "kamp_lite"),
                         labels = c("K", "Kinhom", "perm", "KAMP", "KAMP lite")),
         term = factor(term, levels = c("doc", "p_immune", "age", "stage"),
                       labels = c("Degree of Colocalization at r = 50", "Proportion of immune cells",
                                  "Age", "Cancer stage"))) %>%
  filter(method != "KAMP lite", method != "perm") %>%
  ggplot(aes(x = method, y = estimate, color = method)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_text(aes(label = sig), vjust = -1.2, color = "black", size = 6) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "grey40") +
  facet_wrap(~ term, scales = "free_y", nrow = 1) +
  labs(
    x = "Method",
    y = "Hazard Ratio",
    title = "Hazard ratios for all model covariates at r = 50"
  ) +
  scale_color_manual(values = my_colors) +
  theme_minimal()  +
  theme(legend.position = "none") 
  
```


```{r fig_vpData_survival_talk}
pp/r50 + plot_layout(heights = c(1.5, 1))
```


